Wavelength convergence of the ${\mathrm{\:E}\times\mathrm{B}}$ staircase pattern in flux tube simulations of ion temperature gradient driven turbulence

### 2023-02-21 briefcommunicaton/briefcommunication.tex
#37: widehat for FT{wexb}

\newcommand{\wexb}{\omega_{\mathrm{\:E \times B}}}
\newcommand{\exb}{\mathrm{\:E \times B}}
\newcommand{\hatwexbkvec}{\hat{\omega}_{\mathrm{\:E \times B},\mathbf{k}}}
\newcommand{\NR}{N_\mathrm{R}}
\newcommand{\NB}{N_\mathrm{B}}
\newcommand{\vthref}{v_\mathrm{th,ref}}
\newcommand{\Rref}{R_\mathrm{ref}}
\newcommand{\rlt}{R/L_T}
\newcommand{\rhothref}{\rho_\mathrm{th,ref}}

@@

%%% DIAGNOSTICS ================================================================================================================
The ${\mathrm{\:E}\times\mathrm{B}}$ staircase pattern is manifest as radial structure formation in the ${\mathrm{\:E}\times\mathrm{B}}$ shearing rate defined by\cite{rath2016, doi:10.1063/1.3005380, rath2016, peeters2016}
\begin{equation}
	\omega_{\mathrm{\:E \times B}} = \frac{1}{2} \frac{\partial^2 \langle \Phi \rangle}{\partial \psi^2},
	\label{eq:shearingrate}
\end{equation}
\increasecounter{eq}{1}{1}
where $\langle \Phi \rangle$ is the zonal electrostatic potential normalized by $\rho_\ast T_\mathrm{ref}/e$ ($\rho_\ast = \rho_\mathrm{th,ref}/R_\mathrm{ref}$ is the reference thermal Larmor radius normalized with the reference major radius $R_\mathrm{ref}$, $T_\mathrm{ref}$ is a reference temperature, $e$ is the elementary charge) and $\psi$ is the radial coordinate that labels the flux surfaces normalized by $\rho_\mathrm{th,ref}$.
The E$\times$B shearing rate $\omega_{\mathrm{\:E \times B}}$ is the radial derivative of the advecting zonal flow velocity \cite{doi:10.1063/1.871313, doi:10.1063/1.872847} and quantifies the zonal flow induced shearing of turbulent structures \cite{doi:10.1063/1.859529, doi:10.1063/1.871313, doi:10.1063/1.872367}. 
Here, the zonal potential is estimated from the electrostatic potential $\phi$ on the two dimensional $\psi$-$\zeta$-plane at the low field side according to
\begin{equation}
\langle \Phi \rangle = \frac{1}{l_\zeta} \int_0^{l_\zeta} \mathrm{d}\zeta ~ \phi(\psi,\zeta,s=0).
\end{equation}
In order to diagnose the temporal evolution of the staircase pattern and to obtain an estimate of its amplitude the radial Fourier transform of the ${\mathrm{\:E}\times\mathrm{B}}$ shearing rate is considered. 
It is defined by
\begin{equation}
	\omega_{\mathrm{\:E \times B}} = \sum_{k_j} \hat \omega_{\mathrm{\:E \times B}}(k_j,t) \times \exp(\mathrm{i} k_j \psi),
	\label{eq:shearingrate_fourier}
\end{equation}
\increasecounter{eq}{1}{1}
were $\hat \omega_{\mathrm{\:E \times B}}$ is the complex Fourier coefficient and 
% use j as zonal flow mode number to distinguish it from the imaginary unit i
\begin{equation}
	k_\mathrm{j} = \frac{2\pi \mathrm{j}}{l_\psi}
\end{equation}
\increasecounter{eq}{1}{1}
% explicitly introduce the terms zonal flow wave vector and zonal flow mode number for later usage
defines the zonal flow wave vector with the zonal flow mode number $j$ ranging in $-(N_\psi -1)/2 \leq j \leq (N_\psi -1)/2 $.
% define the zonal flow modes' amplitude in terms of the shearing rate
Based on the defintions above, the shear carried by the zonal flow mode with wave vector $k_j$ is defined by $|\widehat{\omega}_{\mathrm{\:E\times B}}|_{k_\mathrm{j}} = 2 |\hat \omega_{\mathrm{\:E \times B}}|$. 
% introduces the term basic mode of the pattern for later usage
In general, the zonal flow mode that dominates the ${\mathrm{\:E}\times\mathrm{B}}$ staircase pattern, \textemdash also referred to as the \textit{basic mode } of the pattern in this work \textemdash exhibits the maximum amplitude in the spectrum $|\widehat{\omega}_{\mathrm{\:E\times B}}|_{k_\mathrm{j}}$.  
Consistent with Ref.~\cite{peeters2016} the turbulence level is quantified by the turbulent heat conduction coefficient $\chi$, which is normalized by $\rho_\mathrm{th,ref}^2 v_{\mathrm{th,ref}}/R_\mathrm{ref}$ ($v_{\mathrm{th,ref}} = \sqrt{2 T_\mathrm{ref}/m_\mathrm{ref}}$ is the reference thermal velocity and $m_\mathrm{ref}$ is the reference mass) throughout this work. \bigskip

>>>>

\newcommand{\wexb}{\omega_{\mathrm{\:E \times B}}}
\newcommand{\exb}{\mathrm{\:E \times B}}
\newcommand{\hatwexbkvec}{\widehat{\omega}_{\mathrm{\:E \times B},\mathbf{k}}}
\newcommand{\NR}{N_\mathrm{R}}
\newcommand{\NB}{N_\mathrm{B}}
\newcommand{\vthref}{v_\mathrm{th,ref}}
\newcommand{\Rref}{R_\mathrm{ref}}
\newcommand{\rlt}{R/L_T}
\newcommand{\rhothref}{\rho_\mathrm{th,ref}}

@@

%%% DIAGNOSTICS ================================================================================================================
The ${\mathrm{\:E}\times\mathrm{B}}$ staircase pattern is manifest as radial structure formation in the ${\mathrm{\:E}\times\mathrm{B}}$ shearing rate defined by\cite{rath2016, doi:10.1063/1.3005380, rath2016, peeters2016}
\begin{equation}
	\omega_{\mathrm{\:E \times B}} = \frac{1}{2} \frac{\partial^2 \langle \Phi \rangle}{\partial \psi^2},
	\label{eq:shearingrate}
\end{equation}
\increasecounter{eq}{1}{1}
where $\langle \Phi \rangle$ is the zonal electrostatic potential normalized by $\rho_\ast T_\mathrm{ref}/e$ ($\rho_\ast = \rho_\mathrm{th,ref}/R_\mathrm{ref}$ is the reference thermal Larmor radius normalized with the reference major radius $R_\mathrm{ref}$, $T_\mathrm{ref}$ is a reference temperature, $e$ is the elementary charge) and $\psi$ is the radial coordinate that labels the flux surfaces normalized by $\rho_\mathrm{th,ref}$.
The E$\times$B shearing rate $\omega_{\mathrm{\:E \times B}}$ is the radial derivative of the advecting zonal flow velocity \cite{doi:10.1063/1.871313, doi:10.1063/1.872847} and quantifies the zonal flow induced shearing of turbulent structures \cite{doi:10.1063/1.859529, doi:10.1063/1.871313, doi:10.1063/1.872367}. 
Here, the zonal potential is estimated from the electrostatic potential $\phi$ on the two dimensional $\psi$-$\zeta$-plane at the low field side according to
\begin{equation}
\langle \Phi \rangle = \frac{1}{l_\zeta} \int_0^{l_\zeta} \mathrm{d}\zeta ~ \phi(\psi,\zeta,s=0).
\end{equation}
In order to diagnose the temporal evolution of the staircase pattern and to obtain an estimate of its amplitude the radial Fourier transform of the ${\mathrm{\:E}\times\mathrm{B}}$ shearing rate is considered. 
It is defined by
\begin{equation}
	\omega_{\mathrm{\:E \times B}} = \sum_{k_j} \widehat \omega_{\mathrm{\:E \times B}}(k_j,t) \times \exp(\mathrm{i} k_j \psi),
	\label{eq:shearingrate_fourier}
\end{equation}
\increasecounter{eq}{1}{1}
were $\widehat \omega_{\mathrm{\:E \times B}}$ is the complex Fourier coefficient and 
% use j as zonal flow mode number to distinguish it from the imaginary unit i
\begin{equation}
	k_\mathrm{j} = \frac{2\pi \mathrm{j}}{l_\psi}
\end{equation}
\increasecounter{eq}{1}{1}
% explicitly introduce the terms zonal flow wave vector and zonal flow mode number for later usage
defines the zonal flow wave vector with the zonal flow mode number $j$ ranging in $-(N_\psi -1)/2 \leq j \leq (N_\psi -1)/2 $.
% define the zonal flow modes' amplitude in terms of the shearing rate
Based on the defintions above, the shear carried by the zonal flow mode with wave vector $k_j$ is defined by $|\widehat{\omega}_{\mathrm{\:E\times B}}|_{k_\mathrm{j}} = 2 |\widehat \omega_{\mathrm{\:E \times B}}|$. 
% introduces the term basic mode of the pattern for later usage
In general, the zonal flow mode that dominates the ${\mathrm{\:E}\times\mathrm{B}}$ staircase pattern, \textemdash also referred to as the \textit{basic mode } of the pattern in this work \textemdash exhibits the maximum amplitude in the spectrum $|\widehat{\omega}_{\mathrm{\:E\times B}}|_{k_\mathrm{j}}$.  
Consistent with Ref.~\cite{peeters2016} the turbulence level is quantified by the turbulent heat conduction coefficient $\chi$, which is normalized by $\rho_\mathrm{th,ref}^2 v_{\mathrm{th,ref}}/R_\mathrm{ref}$ ($v_{\mathrm{th,ref}} = \sqrt{2 T_\mathrm{ref}/m_\mathrm{ref}}$ is the reference thermal velocity and $m_\mathrm{ref}$ is the reference mass) throughout this work. \bigskip
